Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields

### 2017-01-27 Master Thesis/synopsis.tex
Add Theorem about unramified extensions

We denote the valuation ring of $v\in\operatorname{Spv}(K)$ by $(\mathcal{O}_v,\mathfrak{m}_v)$ and set $U_v=\mathcal{O}_v^\times$ and $k(v)=\mathcal{O}_v/\mathfrak{m}_v$. The \textit{rank}\index{valuation!rank} of $v$ is defined as the Krull dimension of its valuation ring $\mathcal{O}_v$. The \textit{rational rank}\index{valuation!rational rank} of $v:K\to\Gamma_v\cup\{\infty\}$ is defined as the rank of its valuation group $\Gamma_v$ as a $\mathbb{Z}$-module:
\[\operatorname{rrank}(v) = \dim_\mathbb{Q}(\Gamma_v\otimes_\mathbb{Z}\mathbb{Q}) \]
We write $K_v$ for the quotient field of a henselization of $\mathcal{O}_v$.
\end{definition}

@@

@@

\begin{proof}
See Valued Fields.
\end{proof}

@@

Let $v\in\operatorname{Spv}_\text{d}^{(1)}$ and let $N$ be the normal subgroup of $G_{K_v}$ generated by all subgroups $G_{K_w}$ for $w\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ with $v(w) = v$. Since a global field has no non-trivial algebraic extension in which every prime splits completely, we get $N=G_{K_v}$. This shows that $K'_{v'} = (K_v)'$ is a henselization of $K'$ with respect to $v'=\sigma(v)$. Furthermore, let $w_1,w_2\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ with $w_1\neq w_2$ and $v=v(w_1)=v(w_2)$, then:
\[ \sigma(G_{K^\text{nr}_v}) = \sigma(G_{K_{w_1}}\cap G_{K_{w_2}}) = G_{K'_{w'_1}}\cap G_{K'_{w'_2}} = G_{(K')^\text{nr}_{v'}} \]
Therefore, $\sigma$ induces an isomorphism:
\[ G_{k(v)} \cong G_{K_v}/G_{K_v^\text{nr}} \to G_{K'_{v'}}/G_{(K')_{v'}^\text{nr}}\cong G_{k(v')} \]
The rest follows from the Theorem of Neukirch-Uchida.
\end{proof}

>>>>

We denote the valuation ring of $v\in\operatorname{Spv}(K)$ by $(\mathcal{O}_v,\mathfrak{m}_v)$ and set $U_v=\mathcal{O}_v^\times$ and $k(v)=\mathcal{O}_v/\mathfrak{m}_v$. The \textit{rank}\index{valuation!rank} of $v$ is defined as the Krull dimension of its valuation ring $\mathcal{O}_v$. The \textit{rational rank}\index{valuation!rational rank} of $v:K\to\Gamma_v\cup\{\infty\}$ is defined as the rank of its valuation group $\Gamma_v$ as a $\mathbb{Z}$-module:
\[\operatorname{rrank}(v) = \dim_\mathbb{Q}(\Gamma_v\otimes_\mathbb{Z}\mathbb{Q}) \]
We write $K_v$ for a henselization of $K$ with respect to $v$.
\end{definition}

@@

\begin{theorem}\label{theorem-maxunramified}
Let $K$ be a henselian field with respect to $v: K\to\Gamma\cup\{\infty\}$. Let $K^\text{nr}$ be the maximal unramified extension of $K$ with the unique extension $w$ of $v$. Then:
\begin{enumerate}[(i)]
\item $w$ has value group $\Gamma$ and residue class field $k(v)^\text{s}$.
\item $K^\text{nr}$ is a Galois extension of $K$ and $\operatorname{Gal}(K^\text{nr}/K)\cong G_{k(v)}$. There exists an inclusion-preserving one-to-one correspondence: 
\begin{align*}
\{ \text{Intermediate extensions in $K^\text{nr}/K$} \} &\longrightarrow \{\text{Separable extensions of $k(v)$} \}\\ 
L &\longmapsto k(v_L)  
\end{align*}
where $v_L$ is the unique extension of $v$ to $L$.
\end{enumerate}
\end{theorem}

\begin{proof}
See Valued Fields 5.2.7.
\end{proof}

\begin{proof}
See Valued Fields 4.4.1.
\end{proof}

@@

@@

Let $v\in\operatorname{Spv}_\text{d}^{(1)}$ and let $N$ be the normal subgroup of $G_{K_v}$ generated by all subgroups $G_{K_w}$ for $w\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ with $v(w) = v$. Since a global field has no non-trivial algebraic extension in which every prime splits completely, we get $N=G_{K_v}$. This shows that $K'_{v'} = (K_v)'$ is a henselization of $K'$ with respect to $v'=\sigma(v)$. Furthermore, let $w_1,w_2\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ with $w_1\neq w_2$ and $v=v(w_1)=v(w_2)$, then:
\[ \sigma(G_{K^\text{nr}_v}) = \sigma(G_{K_{w_1}}\cap G_{K_{w_2}}) = G_{K'_{w'_1}}\cap G_{K'_{w'_2}} = G_{(K')^\text{nr}_{v'}} \]
Therefore, $\sigma$ induces with Theorem \ref{theorem-maxunramified} an isomorphism:
\[ G_{k(v)} \cong G_{K_v}/G_{K_v^\text{nr}} \to G_{K'_{v'}}/G_{(K')_{v'}^\text{nr}}\cong G_{k(v')} \]
The rest follows from the Theorem of Neukirch-Uchida.
\end{proof}
